---
title: "eds_221_day7_interactive"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---
```{r}
library(tidyverse)
library(palmerpenguins)
library(lubridate)
```

starting off with a data wrangling refresher
Refresher 1: Starting with the penguins dataset in the palmerpenguins package, write a single piped sequence in which you:

Only include penguins at Biscoe and Dream islands
Remove the year and sex variables
Add a new column called body_mass_kg, with penguin mass converted from grams to kilograms
Rename the island variable to location


```{r}
penguins1 <- penguins %>%
  filter(island %in% c("Biscoe", "Dream")) %>%
  select(!c(sex, year)) %>%
  mutate(body_mass_kg = body_mass_g / 1000) %>%
  rename(location = island)
```

Refresher 2: Staring with the penguins dataset in the palmerpenguins package, write a single piped sequence in which you:

Limit to only Adelie penguins
Remove any observations where flipper_length_mm is NA (hint: !is.na())
Group the data by sex
Find the mean (mean()), standard deviation (sd) and sample size (n) of flipper_length_mm for male and female Adelie penguins, returned in a nice summary table

```{r}
penguins2 <- penguins %>%
  filter(species=="Adelie") %>%
  filter(!is.na(flipper_length_mm)) %>%
  group_by(sex) %>%
  summarize("Mean" = mean(flipper_length_mm), "SD" = sd(flipper_length_mm), "Sample Size" = n())
```

moving on to more data wrangling!









































